Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-01-07 doc/invt.tex
add discussion on the tridiagonal matrix

To do so, we first recall
the formal solution of Eq. \eqref{eq:dxdt_alpha}:
%
\begin{equation}
x(t) = x(0) \, e^{-\lambda \, q(t)}
+ \int_0^t \dot u(t') \, \xi(t') \, dt',
\label{eq:xt_solution}
\end{equation}
%
where,
%
\begin{equation}
q(t) \equiv \int_0^t \alpha(t') \, dt',
\label{eq:qt_definition}
\end{equation}
%
and
%
\begin{align}
u(q(t'))
&\equiv
e^{-\lambda \, q(t) + \lambda \, q(t')}.
\label{eq:ut_definition}
\end{align}

We shall further demand that
%
\begin{equation}
  \lim_{t \to \infty} q(t) \to \infty.
  \label{eq:qt_limit}
\end{equation}
%
Then, for a large time $t$,
the first term on the right-hand side
of Eq. \eqref{eq:xt_solution} can be neglected, and
%
\begin{align}
\left\langle x^2(t) \right\rangle
%&=
%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%\notag
%\\
&=
\int_0^t \int_0^t
  \dot u(q(t')) \, \dot u(q(t'')) \,
  \kappa(t' - t'') \, dt'' \, dt'.
\label{eq:x2t_average}
\end{align}

Variating this expression and
using the Euler-Lagrange equation yields
$$
\begin{aligned}
0
&=
\frac{d}{d\tau} \int_0^t \dot u(q(t')) \, \kappa(t' - \tau) \, dt'
= \int_0^t \ddot u(q(t')) \, \kappa(t' - \tau) \, dt',
\end{aligned}
$$
where we have dropped the boundary terms
by using Eq. \eqref{eq:kappat_limit}.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
In order for this to hold for any $\tau$,
we must have
%
\begin{equation}
\ddot u(q(t')) = 0,
\qquad
\mathrm{or}
\;\;
\dot u(q(t')) = c,
\label{eq:ddu_eq_0}
\end{equation}
%
where $c$ is a constant of $t'$.
%
Using Eq. \eqref{eq:ut_definition}
we get
$$
e^{-\lambda \, q(t) + \lambda \, q(t')}
=
c \, (t' + t_0),
$$
where $t_0$ is a constant.
%
Taking the logarithm, and differentiating this with respect to $t'$
yields Eq. \eqref{eq:alpha_opt}.

@@

In analogous to Eq. \eqref{eq:x2t_average},
we can rewrite Eq. \eqref{eq:y2_sum} as
%
\begin{align}
  \mathcal E
  %&= \sum_{i = 2}^n \langle y_i^2 \rangle
  %\notag
  %\\
  &=
  \int_0^t \int_0^t
  \sum_{i = 2}^n
  \dot u_i(q(t')) \,
  \dot u_i(q(t'')) \,
  \kappa_i(t' - t'') \, dt' \, dt'',
  \notag
  %\label{eq:y2_int}
\end{align}
%
where
\begin{align*}
  u_i(q(t'))
  &\equiv
  e^{\lambda_i \, [q(t') - \, q(t)]},
  \\
  \kappa_i(t - t')
  &\equiv
  \left\langle \eta_i(t) \, \eta_i(t') \right\rangle,
\end{align*}
%
and the extremal condition of $\mathcal E$ is given by
%
\begin{align}
\sum_{i=2}^n
u_i'(q(\tau))
\int_0^t
\ddot u_i(q(t')) \, \kappa_i(t' - \tau) \, dt' = 0.
\label{eq:optimal_mbin1}
\end{align}
%
If all $\lambda_i$, hence all $u_i$, are the same,
the above condition can be satisfied
by Eq. \eqref{eq:ddu_eq_0},
which is the single-bin case.
%
The general solution of Eq. \eqref{eq:optimal_mbin1}
is harder to obtain.
%

@@

To proceed, we shall further assume perfect sampling,
%
\begin{equation}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  p_i \, \delta_{ij} \, \delta(t - t').
  \label{eq:zz_perfect}
\end{equation}
%
Thus, from Eqs. \eqref{eq:T_orthonormal_rows} and \eqref{eq:eta_def},
we get
%
\begin{equation}
  \kappa_i(t - t')
  =
  \sum_{j,k = 1}^n
  \frac{ T_{ij} } { p_j }
  \frac{ T_{ik} } { p_k }
  \langle \zeta_j(t) \, \zeta_k(t') \rangle
  =
  \delta(t - t').
  \label{eq:kappa_perfect}
\end{equation}
%
Thus, Eq. \eqref{eq:optimal_mbin1}
is simplified as
%
\begin{align}
  \sum_{i=1}^n
  \ddot u_i(q(\tau)) \, u'_i(q(\tau)) = 0.
  \label{eq:optimal_mbin2}
\end{align}
%
%We can study two limiting cases.
%
For $\tau \ll t$,
the dominant contribution comes from
the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
and
$$
\ddot u_n(\tau) = 0,
\quad
\dot u_n(\tau) = c,
\quad
\mathrm{and}\;
u_n(\tau) = c \, (\tau + t_0),
$$
which means
$$
\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
$$

@@

A more direct, and often more accurate, approach
is to assume the functional form
\begin{equation}
\alpha(\tau) = \frac{1}{\lambda \, (\tau + t_0) },
\label{eq:alpha_trial}
\end{equation}
and find the optimal value of $\lambda$.
%
Using Eqs. \eqref{eq:kappa_perfect}, \eqref{eq:alpha_trial}
in Eq. \eqref{eq:y2_sum}, we get
%
\begin{equation}
\mathcal E
=
\int_0^t
\sum_{i=0}^n \dot u_i^2(q(t')) \, dt'
\stackrel{t \to \infty}
{=\joinrel=\joinrel=}
\frac{1}{t}
\sum_{i = 2}^n
\frac{ \lambda_i^2 } { \lambda \, (2 \, \lambda_i - \lambda) },
\label{eq:y2_sum_perfect_invt}
\end{equation}
%
and the optimal $\lambda$
can be found from the smallest root of
%
\begin{equation}
\sum_{i = 2}^n
\frac{ \lambda_i - \lambda }
{ \left(2 - \lambda/ \lambda_i \right)^2 }
= 0.
\end{equation}

@@

\subsection{Example}

@@

This matrix is clearly symmetric, and hence
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i \equiv 1/n$.
%
It admits $n$ eigenvectors,
$\pmb\theta^{(1)}, \dots, \pmb\theta^{(n)}$,
given by
%
\begin{equation}
\theta^{(k)}_i
= \cos \frac{ (2 \, i - 1) k \pi}{2 \, n}.
\label{eq:wband_eigenvector}
\end{equation}
%
with eigenvalues
\note{To show Eqs.
  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue},
  we first define $m_l = 0$, for $l \le 0$, then
$$
\begin{aligned}
&\sum_{i = 1}^n
  \theta^{(k)}_i
  \, w_{ij}
=
- \theta^{(k)}_j \, m_0
\\
&
+\sum_{i=1}^n
  \theta^{(k)}_i
  \left(
    m_{i - j} + m_{j - i} + m_{i + j - 1} + m_{2n+1-i-j}
  \right)
.
\end{aligned}
$$

Now
$$
\begin{aligned}
  &
  \sum_{i=1}^n
  \theta^{(k)}_i \,
  \left(
    m_{i - j} + m_{2n+1-i-j}
  \right)
  \\
  &=
  \sum_{l=1-j}^{n-j} \theta^{(k)}_{j+l} \, m_l
  +
  \sum_{l=n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, m_l
  \\
  &=
  \sum_{l=1-j}^{2n-j} \theta^{(k)}_{j+l} \, m_l
  =
  \sum_{l=0}^{b} m_l \, \cos\frac{(j+l-\frac12)k\pi}{n},
\end{aligned}
$$
where we have used
$$
\theta^{(k)}_i = \theta^{(k)}_{1-i} = \theta^{(k)}_{2n+1-i}.
$$
Similarly,
$$
\begin{aligned}
  \sum_{i=1}^n
  \theta^{(k)}_i
  \left(
    m_{j - i} + m_{i+j-1}
  \right)
  =
  \sum_{l=0}^{b} m_l \, \cos\frac{(j-l-\frac12)k\pi}{n}.
\end{aligned}
$$
%
So
$$
\begin{aligned}
\sum_{i = 1}^n \theta^{(k)}_i \, w_{ij}
&=
- \theta^{(k)}_j \, m_0
+\sum_{l=0}^{b} m_l \, 2 \, \cos\frac{lk\pi}{n} \, \cos\frac{(j-\frac12)k\pi}{n}
\\
&= \theta^{(k)}_j \, \lambda_k,
\end{aligned}
$$
with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}.
}
%
\begin{equation}
\lambda_k = m_0 + \sum_{l = 1}^b 2 \, m_l \cos \frac{k l \pi}{n}.
\label{eq:wband_eigenvalue}
\end{equation}

@@

@@

2. Multiple-bin update scheme.

>>>>

To do so, we first recall
the formal solution of Eq. \eqref{eq:dxdt_alpha}:
%
\begin{equation}
x(t) = x(0) \, e^{-\lambda \, q(t)}
+ \int_0^t \dot u\bigl( q(t') \bigr) \, \xi(t') \, dt',
\label{eq:xt_solution}
\end{equation}
%
where,
%
\begin{equation}
q(t) \equiv \int_0^t \alpha(t') \, dt',
\label{eq:qt_definition}
\end{equation}
%
and
%
\begin{align}
u(z)
&\equiv
e^{-\lambda \, q(t) + \lambda \, z}.
\label{eq:ut_definition}
\end{align}

We shall further demand that
%
\begin{equation}
  \lim_{t \to \infty} q(t) \to \infty.
  \label{eq:qt_limit}
\end{equation}
%
Then, for a large time $t$,
the first term on the right-hand side
of Eq. \eqref{eq:xt_solution} can be neglected, and
%
\begin{align}
\left\langle x^2(t) \right\rangle
%&=
%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%\notag
%\\
&=
\int_0^t \int_0^t
  \dot u\bigl( q(t') \bigr) \,
  \dot u\bigl( q(t'') \bigr) \,
  \kappa(t' - t'') \, dt'' \, dt'.
\label{eq:x2t_average}
\end{align}

Variating this expression and
using the Euler-Lagrange equation yields
$$
\begin{aligned}
0
&=
\frac{d}{d\tau} \int_0^t
  \dot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt'
= \int_0^t
  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - \tau) \, dt',
\end{aligned}
$$
where we have dropped the boundary terms
by using Eq. \eqref{eq:kappat_limit}.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
In order for this to hold for any $\tau$,
we must have
%
\begin{equation}
\ddot u\bigl( q(t') \bigr) = 0,
\qquad
\mathrm{or}
\;\;
\dot u\bigl( q(t') \bigr) = c,
\label{eq:ddu_eq_0}
\end{equation}
%
where $c$ is a constant of $t'$.
%
Using Eq. \eqref{eq:ut_definition}
we get
$$
e^{-\lambda \, q(t) + \lambda \, q(t')}
=
c \, (t' + t_0),
$$
where $t_0$ is a constant.
%
Taking the logarithm, and differentiating this with respect to $t'$
yields Eq. \eqref{eq:alpha_opt}.

@@

In analogous to Eq. \eqref{eq:x2t_average},
we can rewrite Eq. \eqref{eq:y2_sum} as
%
\begin{align}
  \mathcal E
  %&= \sum_{i = 2}^n \langle y_i^2 \rangle
  %\notag
  %\\
  &=
  \int_0^t \int_0^t
  \sum_{i = 2}^n
  \dot u_i\bigl( q(t') \bigr) \,
  \dot u_i\bigl( q(t'') \bigr) \,
  \kappa_i(t' - t'') \, dt' \, dt'',
  \label{eq:error_mbin}
\end{align}
%
where
\begin{align*}
  u_i(z)
  &\equiv
  e^{\lambda_i \, [q(t') - z]},
  \\
  \kappa_i(t - t')
  &\equiv
  \left\langle
    \eta_i(t) \, \eta_i(t')
  \right\rangle,
\end{align*}
%
and the extremal condition of $\mathcal E$ is given by
%
\begin{align}
\sum_{i=2}^n
\lambda_i \, u_i\bigl( q(\tau) \bigr)
\int_0^t
\ddot u_i\bigl( q(t') \bigr) \, \kappa_i(t' - \tau) \, dt' = 0.
\label{eq:optimal_mbin}
\end{align}
%
If all $\lambda_i$, hence all $u_i$, are the same,
the above condition can be satisfied
by Eq. \eqref{eq:ddu_eq_0},
which is the single-bin case.
%
The general solution of Eq. \eqref{eq:optimal_mbin}
is harder to obtain.

@@

To proceed, we shall further assume perfect sampling,
%
\begin{equation}
  \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  p_i \, \delta_{ij} \, \delta(t - t').
  \label{eq:zz_perfect}
\end{equation}
%
Thus, from Eqs. \eqref{eq:T_orthonormal_rows} and \eqref{eq:eta_def},
we get
%
\begin{equation}
  \kappa_i(t - t')
  =
  \sum_{j,k = 1}^n
  \frac{ T_{ij} } { p_j }
  \frac{ T_{ik} } { p_k }
  \langle \zeta_j(t) \, \zeta_k(t') \rangle
  =
  \delta(t - t').
  \label{eq:kappa_perfect}
\end{equation}
%
Then, Eq. \eqref{eq:error_mbin} is simplified as
%
\begin{align}
  \mathcal E
  &=
  \int_0^t
  \sum_{i = 2}^n
  \dot u_i^2\bigl( q(t') \bigr) \, dt',
  \label{eq:error_mbin_perfect}
\end{align}
%
and Eq. \eqref{eq:optimal_mbin} becomes
%
\begin{align}
  \sum_{i=1}^n
  \lambda_i \, u_i\bigl( q(\tau) \bigr) \, \ddot u_i\bigl( q(\tau) \bigr) = 0.
  \label{eq:optimal_mbin_perfect}
\end{align}
%
\note{If we treat $u_i$ as the position,
  and $\lambda_i$ as the mass,
  then $\lambda_i \, \ddot u_i$ is the force.
  So the left-hand side of Eq. \eqref{eq:optimal_mbin_perfect}
  is the virial.
  %
  Thus this condition states that the total virial is zero.
}

%
%We can study two limiting cases.
%
For $\tau \ll t$,
the dominant contribution comes from
the term(s) with the smallest eigenvalue $\lambda_i = \lambda_n$,
and
$$
\ddot u_n(\tau) = 0,
\quad
\dot u_n(\tau) = c,
\quad
\mathrm{and}\;
u_n(\tau) = c \, (\tau + t_0),
$$
which means
$$
\alpha(\tau) \approx \frac{1}{\lambda_n \, (\tau + t_0)}.
$$

@@

A more direct, and often more accurate, approach
is to assume the functional form
\begin{equation}
\alpha(\tau) = \frac{1}{\lambda \, (\tau + t_0) },
\label{eq:alpha_trial}
\end{equation}
and find the optimal value of $\lambda$.
%
Using Eq. \eqref{eq:alpha_trial}
in Eq. \eqref{eq:error_mbin_perfect} yield
%
\begin{equation}
\mathcal E
\stackrel{t \to \infty}
{=\joinrel=\joinrel=}
\frac{1}{t}
\sum_{i = 2}^n
\frac{ \lambda_i^2 } { \lambda \, (2 \, \lambda_i - \lambda) },
\label{eq:y2_sum_perfect_invt}
\end{equation}
%
and the optimal $\lambda$
can be found from the smallest real root of
%
\begin{equation}
\sum_{i = 2}^n
\frac{ \lambda_i - \lambda }
{ \left(2 - \lambda/ \lambda_i \right)^2 }
= 0.
\label{eq:optimal_lambda_approx}
\end{equation}

@@

\subsection{Uniform band matrix}

@@

This matrix is clearly symmetric, and hence
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i \equiv 1/n$.
%
It admits $n$ eigenvectors,
$\pmb\theta^{(1)}, \dots, \pmb\theta^{(n)}$,
given by
%
\begin{equation}
\theta^{(k)}_i
= \cos \frac{ \left(i - \frac12\right) (k - 1) \, \pi}{n}.
\label{eq:wband_eigenvector}
\end{equation}
%
with eigenvalues
\begin{equation}
  \lambda_k = m_0 + \sum_{l = 1}^b 2 \, m_l \cos \frac{(k - 1)  \, l \pi}{n}.
\label{eq:wband_eigenvalue}
\end{equation}

\note{To show Eqs.
  \eqref{eq:wband_eigenvector} and \eqref{eq:wband_eigenvalue},
  we first define $m_l = 0$, for $l \le 0$, then
$$
\begin{aligned}
&\sum_{i = 1}^n
  \theta^{(k)}_i
  \, w_{ij}
=
- \theta^{(k)}_j \, m_0
\\
&
+\sum_{i=1}^n
  \theta^{(k)}_i
  \left(
    m_{i - j} + m_{j - i} + m_{i + j - 1} + m_{2n+1-i-j}
  \right)
.
\end{aligned}
$$

Now
$$
\begin{aligned}
  &
  \sum_{i=1}^n
  \theta^{(k)}_i \,
  \left(
    m_{i - j} + m_{2n+1-i-j}
  \right)
  \\
  &=
  \sum_{l=1-j}^{n-j} \theta^{(k)}_{j+l} \, m_l
  +
  \sum_{l=n-j+1}^{2n-j} \theta^{(k)}_{2n+1-j-l} \, m_l
  \\
  &=
  \sum_{l=1-j}^{2n-j} \theta^{(k)}_{j+l} \, m_l
  =
  \sum_{l=0}^{b} m_l \, \cos\frac{(j+l-\frac12)(k-1)\pi}{n},
\end{aligned}
$$
where we have used
$$
\theta^{(k)}_i = \theta^{(k)}_{1-i} = \theta^{(k)}_{2n+1-i}.
$$
Similarly,
$$
\begin{aligned}
  \sum_{i=1}^n
  \theta^{(k)}_i
  \left(
    m_{j - i} + m_{i+j-1}
  \right)
  =
  \sum_{l=0}^{b} m_l \, \cos\frac{(j-l-\frac12)(k-1)\pi}{n}.
\end{aligned}
$$
%
So
$$
\begin{aligned}
\sum_{i = 1}^n \theta^{(k)}_i \, w_{ij}
&=
- \theta^{(k)}_j \, m_0
+\sum_{l=0}^{b} m_l \, 2 \, \cos\frac{(k-1)l\pi}{n} \, \cos\frac{(j-\frac12)(k-1)\pi}{n}
\\
&= \theta^{(k)}_j \, \lambda_k,
\end{aligned}
$$
with $\lambda_k$ given by Eq. \eqref{eq:wband_eigenvalue}.
}
%

\subsubsection{Tridiagonal matrix}

Particularly, if $m_2 = \dots = m_b = 0$,
the matrix $\mathbf w$ is tridiagonal,
%
\begin{equation}
\arraycolsep=1.4pt\def\arraystretch{1.4}
\mathbf w
=
\left(
  \begin{array}{cccccccc}
    1 - m_1   & m_1 & 0 & \dots & 0 \\
    m_1 & 1 - 2 \, m_1  & m_1 & \dots & 0 \\
    \vdots & &  & & \vdots \\
    0 & \dots & m_1 & 1 - 2 \, m_1  & m_1 \\
    0 & \dots & 0 & m_1 & 1 - m_1
  \end{array}
\right).
\label{eq:wtridiag}
\end{equation}
%
We have
\begin{equation}
  \lambda_k = 1 - 4 \, m_1 \sin^2 \frac{(k - 1) \, \pi}{2n}.
\label{eq:wtridiag_eigenvalue}
\end{equation}

Thus, this updating matrix requires
$$
m_1 < \frac{1}{4 \, \cos^2\frac{\pi}{2n} }.
$$

For a large $n \gg 1$,
we can convert Eq.
\eqref{eq:y2_sum_perfect_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral,
$$
\begin{aligned}
\mathcal E
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \left(1 - 4 \, m_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, m_1 \, \cos^2 p - \lambda \right) }
\, dp
\notag \\
&=
\frac{n}{4 \, t}
\left(
  \frac{2 - 4 \, m_1 + \lambda}{ \lambda }
  +
  \frac{ \lambda }
  { \sqrt{ (2 - \lambda) (2 - 8 \, m_1 - \lambda) } }
\right).
\end{aligned}
$$
\note{
  The integral is done by contour integration.
  %
$$
\begin{aligned}
\mathcal E
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, m_1 \, \cos^2 p \right)^2 }
{ 2 - 8 \, m_1 \, \cos^2 p - \lambda }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 - m_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
{ 2 - 2 \, m_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\, \frac{dz}{z},
\end{aligned}
$$
where the contour is around the unit circle in the complex plane.
%
This integral has five poles, one at the origin,
which produces a residue $\frac{2 - 4 \, m_1 - \lambda}{4}$
(from series expansion of small $z$),
and the other four at
$$
z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, m_1},
$$
with two of them inside the unit circle:
$$
z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, m_1 - \lambda}}
{\sqrt{8 m_1}}.
$$
Thus,
$$
\begin{aligned}
  \mathcal E
&=
\frac{n}{\lambda \, t}
\left(
 \frac{2 - 4 \, m_1 - \lambda}{4}
 +
 \sum_{z = z_{\pm} }
 \frac{ \left(z^2 - m_1 (1 + z^2)^2 \right)^2 }
 { 2 z^4 (2 - 4 m_1 - \lambda - 4 m_1 z^2) }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, m_1 - \lambda}{4}
 +
 \frac{ \left(1 - m_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
 { 2 - 4 m_1 - \lambda - 4 m_1 z_{\pm}^2 }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, m_1 - \lambda}{4}
 +
 \frac{ (\lambda/2)^2 }
 { \sqrt{(2-\lambda) (2 - 8 m_1 -\lambda)} }
\right).
\end{aligned}
$$
}
%
Minimizing the function yields
%
\begin{equation}
\lambda = \frac{1 - 4 \, m_1} { 1 - 2 \, m_1 },
\label{eq:lambda_tridiag}
\end{equation}
%
and the minimal error is given by
%
\begin{equation}
\mathcal E
=
\frac{n}{t}
\left(
  1+ \frac{2 \, m_1^2}{1-4 \, m_1}
\right),
\label{eq:error_tridiag}
\end{equation}
%
which increases rapidly with $m_1$.

\subsubsection{Gaussian matrix}

Another common updating scheme
in metadynamics is a Gaussian wave.
%
With $b = n - 1 \gg 1$,
we can take the continuous limit:
$x = l \, \Delta x$,
$\Delta x = \pi/n$,
and
$m_l = m(x) \, \Delta x$.
%
Then
Eq. \eqref{eq:wband_eigenvalue}
can be approximated by an integral:
%
\begin{equation}
\lambda_k
=
2 \int_0^{\pi}
m(x) \, \cos\bigl((k-1) \, x\bigr) \, dx,
\label{eq:lambda_int}
\end{equation}
%
with the normalization
%
\begin{equation}
1 = 2 \int_0^{\pi} m(x) \, dx.
\label{eq:mx_normalization}
\end{equation}

Particularly, for the Gaussian wave
%
\begin{equation}
m(x) = \frac{1}{\sqrt{2\pi\sigma^2}}
\exp\left( -\frac{ x^2 } { 2 \, \sigma^2 } \right),
\end{equation}
%
with $\Delta x \ll \sigma \ll 1$,
we can extend the upper limit of the integrals
in Eqs. \eqref{eq:lambda_int}
and \eqref{eq:mx_normalization} to infinity,
and
\begin{equation}
\lambda_k = \exp\left(
  -\frac{ (k-1)^2 \, \sigma^2 } { 2 }
\right).
\label{eq:lambda_Gaussian}
\end{equation}

Owing to singularity,
we can show that the optimal $\lambda$
in Eq. \eqref{eq:alpha_trial}
\begin{equation}
  \lambda \approx 2 \lambda_{n}.
\end{equation}

@@

To determine the optimal schedule,
we shall test a model system
in the following fashion.
%
Initially $\alpha$ is set to $\alpha_0 = 1/n$
for a sufficiently long time.
%
Then we reset the simulation time $t = 0$,
and switch to
$$
\alpha(t) = \min\left\{ \alpha_0, \frac{ 1 } { \lambda \, t } \right\}.
$$

@@

2. Multiple-bin update scheme.
Error vs. $\lambda$.

3. Optimality of the Single-bin update.
Error vs. $m_1$.
